---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
activity <- read.csv("activity.csv", head = TRUE, colClasses =c("integer", "character", "integer"), na.strings="NA")
head(activity)
```

change the format of date from character to date 
```{r, echo=TRUE}
activity$date <- as.Date(activity$date)
## subset and keep only rows with non "NA" steps
activity2 <- subset(activity, !is.na(activity$steps))
```

## What is mean total number of steps taken per day?
create a histogram of the total no of steps taken every day
```{r, echo=TRUE}
sumofsteps <- tapply(activity2$steps, activity2$date, sum, na.rm=TRUE, simplify=T)
sumofsteps <- sumofsteps[!is.na(sumofsteps)]

hist(x=sumofsteps,
     main="Histogram - Number of Steps Taken Each Day",
     xlab="Number of Steps Daily",
     ylab="Frequency",
     col="blue",
     breaks=10)
```
calculate and report the mean and median no of steps taken per day
```{r, echo=TRUE}
mean(sumofsteps)
median(sumofsteps)
```

## What is the average daily activity pattern?
calculate and report the mean and median no of steps taken per day
```{r, echo=TRUE}
avestep <- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify=T)
avestep2 <- data.frame(interval=as.integer(names(avestep)), ave=avestep)

with(avestep2,
     plot(interval,
          ave,
          type="l",
          xlab="intervals",
          ylab="average num of steps"))

## max no of steps across the intervals
maxstep <- max(avestep2$ave)
avestep2[avestep2$ave == maxstep, ]
```

## Imputing missing values
count no of rows with missing values
```{r, echo=TRUE}
sum(is.na(activity$steps))
```
impute missing values with the mean for the interval
```{r, echo=TRUE}
impute <- activity
nastep <- is.na(impute$steps)
intave <- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify=T)
impute$steps[nastep] <- intave[as.character(impute$interval[nastep])]
```
create a histogram of the total no of steps taken every day, blanks are imputed with mean for the interval
```{r, echo=TRUE}
sumofsteps2 <- tapply(impute$steps, impute$date, sum, na.rm=TRUE, simplify=T)

hist(x=sumofsteps2,
     main="Histogram - Number of Steps Taken Each Day, imputed for NA",
     xlab="Number of Steps Daily",
     ylab="Frequency",
     col="blue",
     breaks=10)
```

create a histogram of the total no of steps taken every day, blanks are imputed with mean for the interval
```{r, echo=TRUE}
sumofsteps2 <- tapply(impute$steps, impute$date, sum, na.rm=TRUE, simplify=T)

hist(x=sumofsteps2,
     main="Histogram -mber of Steps Taken Each Day, imputed for NA",
     xlab="Number of Steps Daily",
     ylab="Frequency",
     col="blue",
     breaks=10)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
weekday <- function(d) {
  wd <- weekdays(d)
  ifelse (wd == "Saturday" | wd == "Sunday", "weekend", "weekday")
}

wx <- sapply(impute$date, weekday)
impute$wk <- as.factor(wx)
head(impute)
```

plot to compare the no of steps on weekdayss and weekend
```{r, echo=TRUE}
wkactivity <- aggregate(steps ~ wk+interval, data=impute, FUN=mean)

library(lattice)
xyplot(steps ~ interval | factor(wk),
       layout = c(1, 2),
       xlab="interval",
       ylab="number of steps",
       type="l",
       lty=1,
       data=wkactivity)
```
